The Sound of Teaching Music: Expert Pianists' Performance Modulations for Novices

### 2021-06-21 paper/manuscript.Rmd
correction from feedback
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

The results from Experiment 2 replicated our earlier findings and provide further evidence that skilled pianists were able to modulate their performance for teaching purposes. As in Experiment 1, we observed slower performance only while teaching the notated articulation. Again, KOR showed that when teaching articulation, participants produced shorter staccato, while there was no significant difference in legato between the two conditions as Experiment 1. While teaching dynamics, KV showed that participants produced louder forte and softer piano. This pattern of exaggeration was more pronounced than in Experiment 1, where only louder forte was observed. Furthermore, we again found that when teaching dynamics, participants made a larger contrast between forte and piano bidirectionally at transition points. Importantly, participants did not modulate performance aspects that were irrelevant with regard to the techniques to be taught (i.e., no modulation of dynamics when teaching articulation and vice versa). Overall, Experiment 2 demonstrated systematic and fine-grained didactic performance modulations for a naturalistic piece of music.

@@

Although participants tended to play slower in the teaching condition than in the performing condition in general, we found a significant difference only for articulation, not for dynamics. As we mentioned in the discussion of Experiment 1, it could be that the metronome beats given prior to each performance discouraged participants from deviating from the prescribed tempo. Another possibility is that expression is actually best taught when leaving the tempo unchanged when it is irrelevant. This would imply that general didactic modulations like slowing down might be less useful in the context of teaching expressive artistic techniques where timing itself can be used to add an expression to a performance.

>>>>

The results from Experiment 2 replicate our earlier findings and provide further evidence that skilled pianists were able to modulate their performance for teaching purposes. As in Experiment 1, we observed slower performance only while teaching the notated articulation. Again, KOR showed that when teaching articulation, participants produced shorter staccato, while there was no significant difference in legato between the two conditions as Experiment 1. While teaching dynamics, KV showed that participants produced louder forte and softer piano. This pattern of exaggeration was more pronounced than in Experiment 1, where only louder forte was observed. Furthermore, we again found that when teaching dynamics, participants made a larger contrast between forte and piano bidirectionally at transition points. Importantly, participants did not modulate performance aspects that were irrelevant with regard to the techniques to be taught (i.e., no modulation of dynamics when teaching articulation and vice versa). Overall, Experiment 2 demonstrated systematic and fine-grained didactic performance modulations for a naturalistic piece of music.

@@

Although participants tended to play slower in the teaching condition than in the performing condition in general, we found a significant difference only for articulation, not for dynamics. As mentioned in the discussion of Experiment 1, it could be that the metronome beats given prior to each performance discouraged participants from deviating from the prescribed tempo. Another possibility is that expression is actually best taught when leaving the tempo unchanged when it is irrelevant. This would imply that general didactic modulations like slowing down might be less useful in the context of teaching expressive techniques where timing itself can be used to add an expression to a performance.
